---
title: "Socioeconomic Deprivation in Germany 2014"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
    toc: no
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning =  FALSE, cache=TRUE)
```

[<back](index)



# Overview 

This is a short example how to use the provided data of [GISD](https://lekroll.github.io/GISD/). We will download the Index and a shapefile for German administrative boundaries at 31.12.2014 form [www.geodatenzentrum.de](http://www.geodatenzentrum.de) to show how a simple analysis can be performed. All analysis is done in R.


## Preparations

To perform the analysis, we need to utilize some packages.
```{r libs, message = FALSE, warning =  FALSE}

# Libraries
require(leaflet) # interactive maps
require(rgdal)   # Read Shapefile
require(readr)  # Read CSV
require(dplyr)  # Datamanagement
require(ggplot)  # Plots
require(ggmap) # Mapping tools
require(stargazer) # Nice Tables
require(RColorBrewer)
require(raster)
require(classInt)

```

As a sample use case for the German Deprivation Index, we are going to Download the Values for 2011 and match them to another dataset containing Rates for High Speed Internet Access on the district (NUTS-3) level. Those Rates represent the Fraction of Households, having Access to high speed Internet in a District. 

### Spatial Polygons

To Plot the data as a map, we also need a spatial boundary file, containting the district boundaries in 2014. This can be obtained easily from [geodatenzentrum.de](http://www.geodatenzentrum.de) (Not shown).

```{r dataspatial, include=F, echo=F, message=FALSE, warning=FALSE, comment=NA}
# Load Spatial Boundaries
temp <- tempfile()
mytempdir <- tempdir()
download.file("http://www.geodatenzentrum.de/auftrag1/archiv/vektor/vg1000_ebenen/2014/vg1000-ew_2014-12-31.geo84.shape.ebene.zip",temp)
unzip(temp, exdir = mytempdir,overwrite = TRUE)
SHPKreise  <- readOGR(paste0(mytempdir,"/vg1000-ew_2014-12-31.geo84.shape.ebenen","/vg1000-ew_ebenen"),
                      "VG1000_KRS")
SHPKreise@data$Kreiskennziffer <- as.integer(as.character(SHPKreise@data$RS))

```

### Data on Socioeconomic Deprivation (GISD)

GISD ist stored on GitHUB you can access this tutorial and all relevant datasets on the [GitHub Page](https://lekroll.github.io/GISD/) for this Project. We are using here data from the **2018 revision** of the score, the release notes can be found [here](https://lekroll.github.io/GISD/Update_2018). We then go to *Files of the 2018 Update->Bund->Kreise) to get the Download link for the Values of the german Kreise (NUTS-3) for 2014.
\ 

```{r datadeprivation,  message = FALSE, warning =  FALSE}

# Deprivation 2014
download.file("https://raw.githubusercontent.com/lekroll/GISD/master/Revisions/2018/Bund/Kreis/Kreis_2014.csv", temp)
Deprivation2014 <- read_csv(temp , locale = locale(encoding = "WINDOWS-1252"))

```

### Outcome: High Speed Internet Access 

\ 
No analysis is complete without an outcome. We choose freely available data from the report https://www.gut-leben-in-deutschland.de which is also hosted freely available on [GitHub](https://github.com/gut-leben-in-deutschland/bericht). The indicator we are using is showing the proportion of households that have access to High Speed Internet (>50 MBit/s).
\ 

```{r datainternet,  message = FALSE, warning =  FALSE}

# High Speed Internet Access
download.file("https://raw.githubusercontent.com/gut-leben-in-deutschland/bericht/master/content/07/03/districts.csv", temp)
HighSpeedInternet <- read_csv(temp)
HighSpeedInternet <- HighSpeedInternet %>% filter(year==2015) %>% dplyr::select(krs,value)
PlotDaten <- merge(Deprivation2014,HighSpeedInternet, by.x="Kreiskennziffer", by.y="krs")
names(PlotDaten)[9]<- "HighSpeedInternet"
SHPKreise@data <- left_join(SHPKreise@data,PlotDaten,by="Kreiskennziffer")

```

## Results

\ 
Now, that we have all the data imported, we can start plotting them. First a map of Deprivation:
\ 

#### Map: Deprivation 2014

```{r mapdeprivation, echo=F, message = FALSE, warning =  FALSE, caption="Deprivation 2014"}

# cf. https://cengel.github.io/rspatial/4_Mapping.nb.html
pal_fun <- colorQuantile("YlOrRd", NULL, n = 5)
breaks_qt <- classIntervals(SHPKreise@data$GISD_Score, n = 5, style = "quantile" )
bbox<- make_bbox(lon = SHPKreise@bbox[1,] , lat = SHPKreise@bbox[2,], f=0)
popup <- paste0("<strong>Deprivationsscore 2014: </strong>", format(SHPKreise@data$GISD_Score, digits = 3))
leaflet(SHPKreise) %>% addProviderTiles("Esri.WorldGrayCanvas") %>%
  addPolygons(data=SHPKreise,
    stroke = FALSE, # remove polygon borders
    fillColor = ~pal_fun(GISD_Score), # set fill color with function from above and value
    fillOpacity = 0.7, smoothFactor = 0.5) %>%
  addLegend("bottomright", 
            colors = brewer.pal(5, "YlOrRd"), 
            labels = paste0("up to ", format(breaks_qt$brks[-1], digits=2)),
            title = 'GISD 2014')


```

\   
\ 

Next a map of High Speed Internet Access:

\  
\ 

#### High Speed Internet Access 2015
 
```{r mapinternet, echo=F, message = FALSE, warning =  FALSE, caption="High Speed Internet Access 2015"}

# cf. https://cengel.github.io/rspatial/4_Mapping.nb.html
pal_fun <- colorQuantile("RdYlGn", NULL, n = 5)
breaks_qt <- classIntervals(SHPKreise@data$HighSpeedInternet, n = 5, style = "quantile" )
bbox<- make_bbox(lon = SHPKreise@bbox[1,] , lat = SHPKreise@bbox[2,], f=0)
popup <- paste0("<strong>High Speed Internet Access 2015: </strong>", format(SHPKreise@data$HighSpeedInternet , digits = 3))
leaflet(SHPKreise) %>% addProviderTiles("Esri.WorldGrayCanvas") %>%
  addPolygons(data=SHPKreise,
    stroke = FALSE, # remove polygon borders
    fillColor = ~pal_fun(HighSpeedInternet), # set fill color with function from above and value
    fillOpacity = 0.7, smoothFactor = 0.5) %>%
  addLegend("bottomright", 
            colors = brewer.pal(5, "RdYlGn"), 
            labels = paste0("up to ", format(breaks_qt$brks[-1]*100, digits=2),"%"),
            title = 'High Speed Internet 2015:')


```
 
\ 
\ 

So, this looks like there's an association. Let's check this out graphically: 

\   
\ 

#### Plot: Association of Deprivation and Internet Access
 
```{r plotassociation, echo=F, message = FALSE, warning =  FALSE}

ggplot(SHPKreise@data, aes(y=HighSpeedInternet*100,x=GISD_Score*100)) + geom_smooth() + coord_fixed() + ylab("High Speed Internet in %") + xlab("GISD Score (rescaled, 0-100)")

```
 
\ 

Well there is one, but is it significant?

\ 

#### Model 1: Linear Regression
 
```{r lmodel, message = FALSE, warning =  FALSE, results='asis'}

linmodel <- lm(SHPKreise$HighSpeedInternet ~ SHPKreise$GISD_Score  )
stargazer(linmodel, type="html")


```
 
\ 

Might be, that we missed some Covariates in our model. What about the size and density of the population in the Regions? Maybe socioeconomic deprivation and urbanity are too closely associated? Let's check this out in a second model:

\ 


#### Model 2: Linear Regression on High Speed Internet Access controlled for population size and density 

```{r lmodel2, message = FALSE, warning =  FALSE, results='asis'}

SHPKreise$ShapeArea  = area(SHPKreise)/1000000
SHPKreise$PopDensity <- SHPKreise$Bevölkerung/SHPKreise$ShapeArea
SHPKreise$Population <- SHPKreise$Bevölkerung
linmodel2 <- lm(SHPKreise$HighSpeedInternet ~ SHPKreise$GISD_Score + SHPKreise$Population + SHPKreise$PopDensity + SHPKreise$ShapeArea)
stargazer(linmodel,linmodel2, type="html")
```

\ 


## Conclusion
 
This example has shown, that it is comparably simple and straightforward to use GISD for Analysis of Regional Inequalities in Germany. It can be used for Analysis regarding the whole population as well as for state specific purposes. If you need help with GISD please [contact us](https://www.rki.de/SharedDocs/Kontaktformulare/K/Kroll_Lars_Eric/Integrator.html).

\ 

## How to cite GISD
Lars Eric Kroll, Maria Schumann, Jens Hoebel, Thomas Lampert.Regional health differences – developing a socioeconomic deprivation index for German. Journal of Health Monitoring 2017 2(2). Robert Koch Institute, Berlin. [Download](http://edoc.rki.de/oa/articles/re4F16l3vDa4g/PDF/20Z7HMEgZo7zk.pdf)

[Import to Reference Manager](http://edoc.rki.de/docviews/endnote.php?id=5130)

\ 
\ 
